An Algorithm for Searching Evolving Graphs in Linear Algebraic Language

### 2016-01-23 paper.tex
Clean up description of Alg 1

Temporal paths connect only active nodes in ways that respect time ordering.
Thus the sequences
$\langle (1, t_1), (1, t_2), (3, t_2), (3, t_3) \rangle$ and
$\langle (1, t_1), (2, t_1), (2, t_3), (3, t_3) \rangle$
are both examples of \textit{temporal paths} from $(1, t_1)$ to $(3, t_3)$,
which are drawn as dotted lines with arrowheads in Figure~\ref{fig:active}.
However, $\langle (1, t_1), (1, t_2), (2, t_2), (3, t_2), (3, t_3) \rangle$
is not a temporal path because node 2 is inactive at time $t_2$.
Similarly,

@@

\begin{defn}
The \textbf{$k$-forward neighbors} of a temporal node $(v, t)$ are the temporal nodes
that are the $k+1$st temporal node in every temporal path of length $k+1$ starting
from $(v, t)$. The \textbf{forward neighbors} of a temporal node $(v, t)$ are its
 \textbf{1-forward neighbors}.
\end{defn}

In Figure~\ref{fig:eg_shortest_path},
the forward-neighbors of $(1, t_1)$ are $(2, t_1)$ and $(1, t_2)$ and
the forward-neighbors of $(2, t_1)$ is $(2, t_3)$.
The 2-forward neighbors of $(1, t_1)$ are
$(2, t_1), (1, t_2), (2, t_2)$ and $(3, t_2)$.
By construction, time stamp of every forward neighbor of an active node $(v, t)$
must be no earlier than $t$.

@@

Note that this notion of distance is not a metric, since the distance from $(v, t)$
to $(w, s)$ will in general differ from the distance of $(v, t)$ from $(w, s)$.

@@

The BFS on evolving graphs is described in Algorithm \ref{alg:bfs}.
Given an evolving graph $G_n$ and a root $(s,t)$,
Algorithm~\ref{alg:bfs} returns all temporal notes reachable from the root and
their distances from the root.
The BFS constructs a tree inductively by discovering all $k$-forward neighbors
of the root before proceeding to all $(k+1)$-forward neighbors of the root.

For Figure \ref{fig:eg_shortest_path}, suppose we start with $(1, t_2)$, then
$frontier$ is $\langle(1,t_2)\rangle$ after step $5$.  Since the only forward-neighbor of
$(1,t_2)$ is $(3,t_2)$, $dict.(3,t_2) =1$ at step $11$ and $next$ is enqueued to be
$\langle (3,t_2) \rangle$ at step $12$. After that $frontier$ is equal to $next$. Since
$frontier$ is not empty, we go through step $7$-$12$ again. This time
the forward-neighbor of $(3,t_2)$ is $(3,t_3)$, so $dict.(3,t_3)=2$ and
$next = \langle (3,t_3) \rangle$ and so on. The loop terminates when $frontier$ is empty.

For each temporal node $(v_i,t_i)$, we use a dictionary $dict.(v_i,t_i)$
to store its distance from $(s, t)$. We also use
two FIFO queues $frontier$ and $next$, where $frontier$
stores the frontier nodes of the current discovered regions and $next$ stores
the frontier nodes of the next level.

\begin{algorithm}[h]
 \SetKwFor{For}{for}{}{end}
 \SetKwFor{While}{while}{}{end}
 \SetKwFor{If}{if}{}{end}
 \SetKwFunction{nei}{forward-neighbors}
 \SetKwFunction{enqueue}{enqueue}
 \SetKwProg{Fn}{function}{}{end}
 \Fn{BFS($G_n, (s,t)$)}{
  $dict.(s,t) = 0$ \\
  $i = 1$ \\
  $fronter = \emptyset$ \\
  \enqueue{$frontier, (s,t)$} \\
 \While{$frontier \ne \emptyset$}{
   $next = \emptyset$ \\
   \For{$(u,t') \in frontier$}{
     \For{$(v,t'') \in$\nei{$(u,t')$}}{
          \If{$(v,t'') \notin d$}{
              $dict.(v,t'') = i$\\
              \enqueue{$next, (v,t'')$}
              }
         }
       }
  $frontier = next$ \\
  $i = i+1$ \\
  }
  \textbf{return} $dict$
}
\caption{Breadth-first search (BFS) on an evolving graph: given
$G_n$, an evolving graph represented by adjacency lists, and a root $(s, t)$.
The algorithm returns $dict$ which stores all reachable node-time stamp pairs from $(s,t)$ such that if $dict.(v_i, t_i) = k$,
the length of the shortest temporal path
from $(s,t)$ to $(v_i, t_i)$ is $k$.
}
\label{alg:bfs}
\end{algorithm}

>>>>

Temporal paths connect only active nodes in ways that respect time ordering.
Thus the sequences
$\langle (1, t_1), (1, t_2), (3, t_2), (3, t_3) \rangle$ and
$\langle (1, t_1), (2, t_1), (2, t_3), (3, t_3) \rangle$
are both examples of \textit{temporal paths} from $(1, t_1)$ to $(3, t_3)$,
which are drawn as dotted lines with arrowheads in Figure~\ref{fig:active}.
However, $\langle (1, t_1), (1, t_2), (2, t_2), (3, t_2), (3, t_3) \rangle$
is not a temporal path because node 2 is inactive at time $t_2$.

@@

\begin{defn}
The \textbf{$k$-forward neighbors} of a temporal node $(v, t)$ are the temporal nodes
that are the $k+1$st temporal node in every temporal path of length $k+1$ starting
from $(v, t)$. The \textbf{forward neighbors} of a temporal node $(v, t)$ are its
1-forward neighbors.
\end{defn}

In Figure~\ref{fig:eg_shortest_path},
the forward neighbors of $(1, t_1)$ are $(2, t_1)$ and $(1, t_2)$ and
the forward neighbors of $(2, t_1)$ is $(2, t_3)$.
The 2-forward neighbors of $(1, t_1)$ are
$(2, t_1), (1, t_2), (2, t_2)$ and $(3, t_2)$.
By construction, time stamp of every forward neighbor of an active node $(v, t)$
must be no earlier than $t$.

@@

Note that this notion of distance is not a metric, since the distance from $(v, t)$
to $(w, s)$ will in general differ from the distance of $(v, t)$ from $(w, s)$
owing to time ordering.

@@

The BFS on evolving graphs is described in Algorithm~\ref{alg:bfs}.
Given an evolving graph $G_n$ and a root $(v_1, t_1)$,
Algorithm~\ref{alg:bfs} returns all temporal notes reachable from the root and
their distances from the root.
$reached$ is a dictionary from temporal nodes to integers whose key set represents
all visited temporal nodes and whose value set are the corresponding distances
from the root.

The BFS constructs a tree inductively by discovering all $k$-forward neighbors
of the root before proceeding to all $(k+1)$-forward neighbors of the root.
Within the outermost loop, the algorithm iterates over $frontier$, a list of all
temporal nodes of distance $k$ from the root. The $nextfrontier$ list is populated
with all temporal nodes that are forward neighbors of any temporal node in the
$frontier$ list which have not yet been reached by the algorithm.

As a simple example, consider the BFS on the example graph in
Figure~\ref{fig:eg_shortest_path} starting from the root $(1, t_2)$.
The $frontier$ list is first initialized to $\{(1,t_2)\}$.
Since the only forward neighbor of $(1,t_2)$ is $(3,t_2)$,
iteration $k=1$ produces $reached[(3,t_2)]=1$ and $nextfrontier = \{(3,t_2)\}$.
In the next iteration $k=2$, the only forward neighbor of $(3,t_2)$ is $(3,t_3)$,
so $reached[(3,t_3)]=2$ and $nextfrontier = \{(3,t_3)\}$ and so on.
The algorithm terminates after $k=3$ after verifying that $(3,t_3)$ has no
forward neighbors.

\begin{algorithm}[h]
 \SetKwFor{For}{for}{}{end}
 \SetKwFor{While}{while}{}{end}
 \SetKwFor{If}{if}{}{end}
 \SetKwFunction{nei}{forwardneighbors}
 \SetKwProg{Fn}{function}{}{end}
 \Fn{BFS($G_n, (v_0,t_0)$)}{
  $reached[(v_0,t_0)] = 0$ \\
  $frontier = \{(v_0,t_0)\}$ \\
  $k = 1$ \\
  \While{$frontier \ne \emptyset$}{
   $nextfrontier = \emptyset$ \\
   \For{$(v,t) \in frontier$}{
     \For{$(v',t') \in$\nei{$(v,t)$}}{
          \If{$(v',t') \notin d$}{
              $reached[(v',t')] = k$\\
              $nextfrontier = nextfrontier \cup \{(v',t')\}$ \\
              }
         }
       }
  $frontier = nextfrontier$ \\
  $k = k+1$ \\
  }
  \textbf{return} $reached$
}
\caption{Breadth-first search (BFS) on an evolving graph $G_n$ starting from a
root $(v_1, t_1)$. The return value, $reached$, is a dictionary mapping all
reachable temporal notes from the root to their distances from the root. At the
end of each iteration $k$, the $frontier$ set contains all temporal nodes of
distance $k$ fro the root.
}
\label{alg:bfs}
\end{algorithm}
